---
title: "Business Intelligence from Web Data Analytics and Data Mining using R and AI - Activity2"
author: "Jannik Guldmand, Sophia Klimova & Adam Dapoz"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: false
    theme: flatly
    highlight: tango
    df_print: paged
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Packages Used}
library(skimr)
library(dplyr)
```


# Introduction
The data we use is comming from a competive tetris ladder https://ch.tetr.io/
The data was webscraped via the api https://ch.tetr.io/api/users/by/league?limit=50&after=23731.327577078657%3A0%3A1e-10

# The Dataset
After scraping we are now ready to look into the data:

## Loading the CSV
```{r loading the data}
#load Tetra League data
data = read.csv("tldata.csv")

head(data)
```





## Preparing the data

```{r}
#make some things factors
data$Country           = as.factor(data$Country)

data$Country = sub("Korea, Republic of", "Republic of Korea", data$Country)
data$Country = sub("Venezuela, Bolivarian Republic of", "Republic of Venezuela", data$Country)
data$Country = sub("Macedonia, the former Yugoslav Republic of", "Republic of Macedonia", data$Country)

data$Rank              = factor(data$Rank, levels=c("D","D+","C-","C","C+","B-","B","B+","A-","A","A+","S-","S","S+","SS","U","X","X+"))

data$Active.This.Week = as.factor(data$Active.This.Week)
data$Active.This.Week = ifelse(data$Active.This.Week == "Yes", 1, 0)

data$Supporter.Status. = as.factor(data$Supporter.Status.)
data$Supporter.Status. = ifelse(data$Supporter.Status. == "Yes", 1, 0)

data$Wins = as.numeric(data$Wins)
data$Games.Played = as.numeric(data$Games.Played)

data$Username = as.character(data$Username)


#remove index (standing does this)
data$X = NULL
```


```{r}
data
```




# Explanatory Data Analysis





## EDA 1.0 Looking into the data


### Dimensions
 
```{r Dimensions }
#Data Dimensions
dim(data) 
```

### Summary

```{r Summary }
#Data Summary
summary(data)
```

```{r Column names }
#list features
names(data)
```


### Structure & data types

```{r Structure & data types}
str(data)
sapply(data, class)
```

### Basic summaries

```{r numeric variables }
# numeric variables
num_vars  <- sapply(data, is.numeric)
summary(data[ , num_vars])   # min–max, quartiles, mean
```

```{r categorical variables}

#drop usernames to avoid this output getting to long
data_no_username = data %>% select(-Username) 

# categorical variables
cat_vars  <- sapply(data_no_username, is.factor) | sapply(data_no_username, is.character)
lapply(data_no_username[ , cat_vars], table)     # frequency tables
```
 
 
```{r}
glimpse(data) 
```


```{r}
skim(data)
```

## EDA 1.1 Missing Value Scan
Missing-value scan
```{r}
colSums(is.na(data))
# Optional: percentage
round(colMeans(is.na(data)) * 100, 2)

```
<hr>

## EDA 1.2 Duplicate rows / keys

```{r}
#duplicates = list()
#k = 1
#
#for (i in 1:(nrow(data) - 1)) {
#  for (j in (i + 1):nrow(data)) {
#    if (data$Username[i] == data$Username[j]) {
#      duplicates[[k]] = c(i, j)
#      k = k + 1
#    }
#  }
#}
```



```{r}
sum(duplicated(data))
# If Username should be unique:
sum(duplicated(data$Username))

#which(duplicated(data$Username))

#data$Username[c(11030, 11178, 11429, 12178, 25056)]
#data[c(11030, 11178, 11429, 12178, 25056), ]
```

```{r}
data %>%
  group_by(Username) %>%
  filter(n() > 1)  
```

```{r}
summary(data)
```



```{r}
#d1 = data[c(which(duplicated(data$Username))), 1]
#d2 = data[c(which(duplicated(data$Username, fromLast=TRUE))), 1]
#
#for (i in d1) for (j in d2) {
#  if (data$Games.Played[i] > data$Games.Played[j])
#    data = data[-j,]
#  else
#    data = data[-i,]
#}

d1 = data[c(which(duplicated(data$Username))), 1]
d2 = data[c(which(duplicated(data$Username, fromLast=TRUE))), 1]
to_remove = integer(0)

for (i in d1) {
  for (j in d2) {
    g1 = data$Games.Played[i]
    g2 = data$Games.Played[j]
    if (data$Username[i] != data$Username[j]) next
    if (g1 > g2) to_remove = c(to_remove, j)
    else if (g1 < g2) to_remove = c(to_remove, i)
  }
}

data = data[-unique(to_remove), ]

data

```

```{r}
data %>%
  group_by(Username) %>%
  filter(n() > 1)  
```
```{r}
sum(duplicated(data))
# If Username should be unique:
sum(duplicated(data$Username))
```


```{r}
data
```


## EDA 1.3 Quick numerical distributions

```{r}
numeric_cols <- names(data)[num_vars]

par(mfrow = c(2, 3))                # 2×3 grid of histograms
for (v in numeric_cols[1:6]) {      # first six just so it’s readable
  hist(data[[v]], main = v, xlab = "")
}
par(mfrow = c(1, 1))

```

<hr>


## EDA 1.4 Plotting

```{r ploting the dataset1 }
plot(data, col=data$RankColour)
```


```{r ploting the dataset2 }
#plot(data$APM, data$Glicko.Rating, data$Winrate, data$PPS)
pairs(data[, c("APM", "Glicko.Rating", "Winrate", "PPS")], col=data$RankColour)
```



```{r ploting the dataset3 }

plot_data = data.frame(data$APM, data$Glicko.Rating, data$Games.Played, data$Winrate, data$PPS)
plot(plot_data, cex=0.3, pch=19, col=data$RankColour)
```

```{r}
# Attacks per minute vs Glicko Rating
plot(data$APM, data$Glicko.Rating, cex=0.3, pch=19, col=data$RankColour)
```
```{r}
# Attacks per minute vs Winrate
plot(data$APM, data$Winrate, cex=0.3, pch=19, col=data$RankColour)
```
 
```{r}
# Attacks per minute vs Winrate
plot(data$Games.Played, data$Winrate, cex=0.3, pch=19, col=data$RankColour)
```
<hr>

## EDA 1.5  Outlier eyeball (boxplots)

```{r}

boxplot(data[ , numeric_cols],
        las = 2,                    # vertical axis labels
        cex.axis = 0.7)             # shrink labels if many vars
```
```{r}
library(tidyr)
library(ggplot2)

data_long <- data %>% 
  pivot_longer(cols = all_of(numeric_cols),
               names_to = "Variable",
               values_to = "Value")

ggplot(data_long, aes(x = Variable, y = Value)) +
  geom_boxplot(outlier.colour = "firebrick") +
  facet_wrap(~ Variable, scales = "free_y", ncol = 4) +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

```
```{r}
#Who has 4.000 wins?????????
which(data$Wins > 3000)
```
```{r}
data[21656,]
```
```{r}
#Who has 8.000 Games played?????????
which(data$Games.Played > 8000) #.... BINGO
```

```{r}
# Who has the APM above 200?
which(data$APM > 200)
```
```{r}
data[c(1,2,4,6),]
```

```{r}
data
```


```{r}
which(data$Glicko.Rating > 4000)
```
```{r}
data[c(1,2),]
```




```{r}
numeric_cols <- names(Filter(is.numeric, data))
#numeric_cols

big_vars <- c("Standing", "Tetra.Rating")
mid_vars <- c("Games.Played", "Wins", "Glicko.Rating")
small_vars <- setdiff(numeric_cols, c(big_vars, mid_vars))

stopifnot(all(big_vars   %in% numeric_cols),
          all(mid_vars   %in% numeric_cols),
          all(small_vars %in% numeric_cols))


par(mfrow = c(1, 3))

boxplot(data[ , big_vars],
        main = "Big scale", las = 2)

boxplot(data[ , mid_vars],
        main = "Mid scale", las = 2)

boxplot(data[ , small_vars],
        main = "Small scale", las = 2)

par(mfrow = c(1, 1))   # reset
```
 
```{r}
## 1. Pick out numeric columns --------------------------------------------
numeric_cols <- names(Filter(is.numeric, data))

## 2. Choose a grid that fits them all (3 columns is usually nice) ---------
nplots <- length(numeric_cols)
ncols  <- 3
nrows  <- ceiling(nplots / ncols)

par(mfrow = c(nrows, ncols),           # grid of plots
    mar = c(4, 4, 2, 0.5))             # tighten margins a bit

## 3. Loop over the variables ---------------------------------------------
for (v in numeric_cols) {
  boxplot(data[[v]],
          main = v,
          ylab = "",                   # leave y-axis label blank
          horizontal = TRUE)           # optional: horizontal boxes
}

par(mfrow = c(1, 1))                   # restore default layout

```
 
 
 
<hr>

## EDA 1.6 Univariate

```{r}
#skewness function
skewness = function(variable) { ( 3*(mean(variable)-median(variable)) ) / sd(variable) }

hist(data$Glicko.Rating)
skewness(data$Glicko.Rating) # 0.1152569
```
### Univariate normality test


```{r}
#data_numerical = data[,c(1,4,5,6,7,8,9,10,11,12,15,17)]

#str(data)

data_numerical = data[,c(1,4,5,6,7,8,9,10,11,12,15)]

for (v in names(data_numerical)) {
  qqnorm(data_numerical[[v]],
         main = v,
         ylab = "Observed Quantiles",
         xlab = "Theoretical Quantiles")
  qqline(data_numerical[[v]], col = "red", lwd = 2)
}
```



<hr>

## EDA 1.7 Multivariate


### Multivariate normality test

```{r}
data_num_complete <- na.omit(data_numerical)   # avoids NA issues
n <- nrow(data_num_complete)
p <- ncol(data_num_complete)
Sx <- cov(data_numerical)
D2 <- mahalanobis(data_numerical, colMeans(data_numerical), Sx)

#Theoretical χ² quantiles
chi_q <- qchisq(ppoints(n, a = 0.5), df = p)

# Chi-square Q-Q plot (multivariate normality)
qqplot(qchisq(ppoints(n, a = 0.5), df = p), D2,
       ylab = "Mahalanobis Distance",
       xlab = bquote("quantiles of " ~ chi[.(p)]^2))

abline(0, 1, col = "red", lwd = 2)

title(main = "Mahalanobis D²", font.main = 2)
```
A comparison of the Mahalanobis D^2 distances with the χ_12^2 reference line shows a pronounced upward curvature and many large outliers, indicating that the joint distribution of the 12 numeric variables deviates substantially from multivariate normality.


### #APM vs PPS 
```{r}
#APM vs PPS
plot(data$PPS, data$APM, cex=0.5, pch=19, col=data$RankColour,
     xlab="Pieces Per Second", ylab="Attack Per Minute",
     main="APM vs PPS, Coloured by Rank")
linearFit = lm(data$APM ~ data$PPS)
abline(linearFit, lwd=2)

summary(linearFit)
par(mfrow=c(2,2))
plot(linearFit, cex=0.5, pch=19, col=data$RankColour)
```

 
 
### Supporter by Rank 
```{r}
#Supporter by Rank
table       = table(data$Supporter.Status., data$Rank)
proportions = proportions(table, margin=2)
#par(mfrow=c(2,1))
barplot(table, xlab="Rank", ylab="Frequency", col=c("#4d4d4d", "orange"),
        main="Frequency of Supporters & Non-Supporters by Rank")
#box()
legend("topright", legend=c("Supporter", "Not Supporter"), col=c("orange", "#4d4d4d"), pch=15)
barplot(proportions, xlab="Rank", ylab="Proportion", col=c("#4d4d4d", "orange"),
        main="Proportion of Supporters & Non-Supporters by Rank")
#box()
```
 
 
 ### Correlation
 
```{r}
#help(cor)

#data$Country = as.factor(data$Country)
data$Rank_number = as.numeric(data$Rank)
#data$Rank_number

#str(data)


data_correlation = cor(data[,c(1,4,5,6,7,8,9,10,11,12,14,15,17)])
data_correlation

#which(abs(data_correlation) > 0.90)


#data[,-c(1,15,16,28,29,43,57.58)]




```
 
 Highly Correlated variables:
 * Standing or Tetra.Rating
 * Games played or Wins
 * APM, PPS, VS, Glicko Rating
 * Rank_number



```{r}
# The ones ADAM think should ve used
linearFit2 = lm(data$Glicko.Rating ~ data$Games.Played + data$APM + data$PPS + data$VS + data$Rating.Deviation + data$Active.This.Week + data$Supporter.Status.)

summary(linearFit2)
```


```{r}
# The ones we keep based on correlation
linearFit3 = lm(data$Glicko.Rating ~ data$Games.Played + data$VS + data$Rating.Deviation + data$Active.This.Week + data$Supporter.Status.)

summary(linearFit3)
```
### Trying to estimate games playe

```{r}
#here we try to predicto or estimate, a players number of games played
#linearFit4 = lm(data$Games.Played ~ data$Standing + data$Winrate + data$APM + data$PPS + data$VS  + data$Glicko.Rating + data$Rating.Deviation + data$Tetra.Rating + data$Active.This.Week + data$Supporter.Status. + data$Rank_number)

linearFit4 = lm(data$Games.Played ~ data$APM + data$PPS + data$VS  + data$Glicko.Rating + data$Rating.Deviation + data$Active.This.Week + data$Supporter.Status.)

summary(linearFit4)
```

```{r}
#Supporter
linearFit5 = lm(data$Supporter.Status. ~ data$APM + data$PPS + data$VS  + data$Glicko.Rating + data$Rating.Deviation + data$Active.This.Week)

summary(linearFit5)
```


<hr>


# Business Intelligence





## BI 2.1 - Ranking UP


```{r}
#boxplots all ranks apm
boxplot(data$APM ~ data$Rank)
```
```{r}
#calculate average stats for each rank
avgAPMs = tapply(data$APM, data$Rank, mean)
avgPPSs = tapply(data$PPS, data$Rank, mean)
avgVSs  = tapply(data$VS,  data$Rank, mean)
par(mfrow=c(1,3))
barplot(avgAPMs, main="avgAPMs by Rank")
barplot(avgPPSs, main="avgPPSs by Rank")
barplot(avgVSs,  main="avgVSs by Rank")
par(mfrow=c(1,1))
```



<hr>


## BI 2.2 - Finding more supporters




<hr>


## BI 2.3 - Maintaining Competitive Integrity

```{r}
#find outlier in B rank
bRanks = data[which(data$Rank == "B"),]
plot(bRanks$Standing, bRanks$APM, cex=1, pch=19, col=bRanks$RankColour,
     xlab="Standing", ylab="Attack Per Minute", main="APM of all Player in B Rank")
data[which(data$Rank == "B" & data$APM > 40),] #users: KAMYI & CRISSELLE
```
