---
title: "TestrisDecisionTree"
author: "Jannik Guldmand"
date: "2025-07-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(rpart)
library(rpart.plot)
```


## Tetris Decision Tree Test


```{r load the tetris data}
data = read.csv("tldata.csv")
```


## Prepare the data (Activity 2)
```{r}
## Preparing the data

#make some things factors
data$Country           = as.factor(data$Country)

data$Country = sub("Korea, Republic of", "Republic of Korea", data$Country)
data$Country = sub("Venezuela, Bolivarian Republic of", "Republic of Venezuela", data$Country)
data$Country = sub("Macedonia, the former Yugoslav Republic of", "Republic of Macedonia", data$Country)
data$Country           = as.factor(data$Country)

data$Rank              = factor(data$Rank, levels=c("D","D+","C-","C","C+","B-","B","B+","A-","A","A+","S-","S","S+","SS","U","X","X+"))

data$Active.This.Week = as.factor(data$Active.This.Week)
data$Active.This.Week = ifelse(data$Active.This.Week == "Yes", 1, 0)

data$Supporter.Status. = as.factor(data$Supporter.Status.)
data$Supporter.Status. = ifelse(data$Supporter.Status. == "Yes", 1, 0)

data$Wins = as.numeric(data$Wins)
data$Games.Played = as.numeric(data$Games.Played)

data$Username = as.character(data$Username)


#remove index (standing does this)
data$X = NULL
str(data)
```

## splitting into test and train
```{r}
set.seed(4)
num_of_splits = 10
splits = sample( rep(1:num_of_splits, ceiling(nrow(data)/num_of_splits) ), nrow(data) )
#splits
```

## put data into test and train
```{r}
train = data[splits!=1,]
test = data[splits==1,]
```

```{r}
nrow(test)
nrow(train)
```



## Custom prepare data for decisiontree
```{r}
#Train: drop username, country
train_without_username = train
train_without_username$Username = NULL
train_without_username$Country = NULL
train_without_username$RankColour = NULL

#Test: drop username, country
test_without_username = test
test_without_username$Username = NULL
test_without_username$Country = NULL
test_without_username$RankColour = NULL
```

```{r}
nrow(test_without_username)
nrow(train_without_username)
```



```{r}
train_without_username
```
```{r}
test_without_username
```


## CART Tetris Decision Tree (TRAIN)
Not using entroy but maximizing split difference 

```{r}
cartfittrain = rpart(Supporter.Status.~., dat=train_without_username[train_without_username$Standing<400,], method="class", control=rpart.control(minsplit=4, cp=0.03))
rpart.plot(cartfittrain, type=2)
```



```{r}
test_result = predict(cartfittrain, newdata=test_without_username, type="class")
test_result
```

```{r}
table(test_without_username$Supporter.Status., test_result)
```
### Lets try to balance the data

```{r}
#make training set with higher proportion of supporters (50/50 split)
allsupporters = data[data$Supporter.Status. == 1,]
allnonsupporters = data[data$Supporter.Status. == 0,]

allsupporters
allnonsupporters

# balanced 1 ---------------------------------------------------------------------------
train_balanced = rbind( allsupporters, 
               allnonsupporters[sample(1:nrow(allnonsupporters), nrow(allsupporters)),] )
#shuffle it
train_balanced = train_balanced[sample(1:nrow(train_balanced), nrow(train_balanced)),]

# balanced 2 ---------------------------------------------------------------------------

train_balanced2 = rbind( allsupporters, 
               allnonsupporters[sample(1:nrow(allnonsupporters), 3*nrow(allsupporters)),] )
#shuffle it
train_balanced2 = train_balanced2[sample(1:nrow(train_balanced2), nrow(train_balanced2)),]


```


```{r}
nrow(train)
```
```{r}
nrow(train_balanced)
length(which(train_balanced$Supporter.Status. == 1))
length(which(train_balanced$Supporter.Status. == 0))

train_balanced
```



#Train: drop username, country
```{r}
train_balanced_without_username = train_balanced
train_balanced_without_username$Username = NULL
train_balanced_without_username$Country = NULL
train_balanced_without_username$RankColour = NULL
```

```{r}
#boxplot(train_balanced_without_username$Supporter.Status.)
hist(train_balanced_without_username$Supporter.Status.)

length(which(train_balanced_without_username$Supporter.Status. == 1))
length(which(train_balanced_without_username$Supporter.Status. == 0))
```




```{r}
cartfittrain_balanced = rpart(Supporter.Status.~., dat=train_balanced_without_username[train_balanced_without_username$Standing<400,], method="class", control=rpart.control(minsplit=4, cp=0.03))
rpart.plot(cartfittrain_balanced, type=2)
```
```{r}
test_balanced_result = predict(cartfittrain, newdata=train_balanced_without_username, type="class")
test_balanced_result
```


```{r}
table(train_balanced_without_username$Supporter.Status., test_balanced_result)
```
```{r}
nrow(train_balanced2)
length(which(train_balanced2$Supporter.Status. == 1))
length(which(train_balanced2$Supporter.Status. == 0))

train_balanced2
```
#Train balanced2: drop username, country
```{r}
train_balanced2_without_username = train_balanced2
train_balanced2_without_username$Username = NULL
train_balanced2_without_username$Country = NULL
train_balanced2_without_username$RankColour = NULL

#YOU DONT TEST ON BALANCED

```


```{r}
cartfittrain_balanced2 = rpart(Supporter.Status.~., dat=train_balanced2_without_username, method="class", control=rpart.control(minsplit=4, cp=0.03))
rpart.plot(cartfittrain_balanced2, type=2)
```
```{r}
summary(cartfittrain_balanced2)
```



```{r}
test_balanced2_result = predict(cartfittrain_balanced2, newdata=test_without_username, type="class")
test_balanced2_result
```
```{r}
nrow(test_balanced2_result)
```
```{r}
nrow(test_without_username)
```



```{r}
a = table(test_without_username$Supporter.Status., test_balanced2_result)
table(test_without_username$Supporter.Status., test_balanced2_result)
```
```{r}
proportions(a,1)
```
```{r}
(3450+44)/3977
```

```{r}
length(which(test_without_username$Supporter.Status. == 1))
length(which(test_without_username$Supporter.Status. == 0))
```
```{r}
train_balanced2_without_username
str(train_balanced2_without_username)

train_balanced3_without_username= train_balanced2_without_username
train_balanced3_without_username$Supporter.Status. = as.factor(train_balanced3_without_username$Supporter.Status.)


train_balanced3_without_username_no_support = train_balanced2_without_username
#train_balanced3_without_username_no_support = train_balanced3_without_username_no_support$supporter.status. = NULL

train_balanced3_without_username_no_support$Supporter.Status. = NULL
train_balanced3_without_username_no_support
```
```{r}
train_balanced3_without_username_no_support
```


## C4.5 Tetris Decision Tree

```{r}
# C4.5 Decision Tree
library(RWeka)
#help(J48)

treeC45 <- J48(train_balanced3_without_username$Supporter.Status.~., data = train_balanced3_without_username)

print(treeC45)
summary(treeC45)
```

```{r}
sink("tree_output.txt")
print(treeC45)
sink()
```

```{r}
test_balanced2_C45result = predict(treeC45, newdata=test_without_username, type="class")
test_balanced2_C45result
```


```{r}
table(test_without_username$Supporter.Status., test_balanced2_C45result)
```
```{r}
#C45 Accuracy
(3339+63)/3977
```



## C5.0 Tetris Decision Tree
```{r}
#install.packages("C50")
library(C50)
#help(C50)
#dattree50 = C5.0(x=dat[,c(1,2,5,6, 9:13)], y=dat$income)
treeC50 = C5.0(x=train_balanced3_without_username_no_support, y=train_balanced3_without_username$Supporter.Status)
```

```{r}
plot(treeC50)
```
```{r}
treeC50
```
```{r}
summary(treeC50)
```



```{r}
test_balanced2_C50result = predict(treeC50, newdata=test_without_username, type="class")
test_balanced2_C50result
```


```{r}
summary(test_balanced2_C50result)
```

```{r}
table(test_without_username$Supporter.Status., test_balanced2_C50result)
```
```{r}
#C50 Accuracy
(3451+44)/3977
```

